Transient dynamics of granular piles: continuum v discrete modelling

### 2016-07-14 paper.tex
Reformatted subfigures

The DEM sample was composed of $\sim13000$ disks with a uniform distribution of 
diameters by volume fractions ($d_{max} = 1.5 d_{min}$). The mean grain 
diameter and mass are $d\simeq 2.455 $~\si{\mm} and $m\simeq 0.0123$~\si{\kg}, 
respectively. The grains are first poured uniformly into a rectangular box of 
given width and then the right-hand side wall is shifted far to the right 
to allow the grains to spread. A stable granular slope of 13.2~\si{\degree} is 
obtained when all grains come to rest; see~\cref{fig:slope_configuration}. This 
procedure leads to a mean packing fraction $\simeq 0.82$. Soil grains with a 
mean density of 2600~\si{\kg\per\m\cubed} and internal friction coefficient of 
0.4 between grains are considered.

\begin{figure}[tbhp]
  \includegraphics[width=0.95\textwidth]{figs/slope_configuration}
  \caption{Initial geometry and dimensions of the pile subjected to a horizontal 
  excitation.}
  \label{fig:slope_configuration}
\end{figure}

@@

For grain-scale simulations, classical DEM approach was used.

@@

\begin{figure}[tbhp]
  \centering
  \includegraphics[width=0.95\textwidth]{figs/mpm_velocity_slope}
  \caption{Initial configuration for the MPM simulation of a pile subjected to 
  horizontal velocity of 50J.}
  \label{fig:mpm_slope_setup}
\end{figure}

@@

\Cref{fig:Gradient_Slope_Profile_200J} shows the initial evolution of the 
granular slope subjected to an initial horizontal energy $E_0 = 61$ (in 
dimensionless units) using MPM. As the granular slope is sheared along the 
bottom, the shear propagates to the top leaving a cavity in the vicinity of the 
left wall. This cavity gets partially filled as the granular mass at the top 
collapse behind the flowing mass due to inertia. Similar behaviour is observed 
during the initial stages of the flow evolution using CD technique 
(\cref{fig:Gradient_Slope_CD_200J}). Due to inertia, the grains at the top 
of the granular heap roll down to fill the cavity, while the pile continues to 
spread.

\begin{figure}[tbph]
  \centering
  \includegraphics[width=\textwidth]{figs/Gradient_Slope_Profile_200J}
  \caption{MPM simulation of the initial stages of granular pile subjected to a 
  gradient horizontal energy.}
  \label{fig:Gradient_Slope_Profile_200J}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{figs/Gradient_Slope_CD_200J}
  \caption{CD simulation of the initial stages of granular pile subjected to a 
  gradient horizontal energy.~\citep{Mutabaruka2013}.}
  \label{fig:Gradient_Slope_CD_200J}
\end{figure}

@@

\begin{figure}[tbph]
  \centering
  \begin{subfigure}[b]{0.9\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/Runout_Eo_MPM_CD_DEM}
    \caption{Run-out distance as a function of normalised input kinetic energy.}
    \label{fig:Runout_Eo_MPM_CD_DEM}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{0.9\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/Tf_vs_Eo_Slope}
    \caption{Duration of run-out as a function of normalised input kinetic energy.}
    \label{fig:Tf_vs_Eo_Slope}
  \end{subfigure}
  \caption{Evolution of run-out and time as a function of the normalised input 
  energy for a pile subjected a gradient horizontal energy.}
  \label{fig:Slope}
\end{figure}

@@

\begin{figure}[tbhp]
  \centering
  \begin{subfigure}[b]{0.9\textwidth}
    \includegraphics[width=\textwidth]{figs/Energy_Slope}
    \caption{Evolution of total kinetic energy with time.}
    \label{fig:energy_slope}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{0.9\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/Normalised_Energy_Time_Slope}
    \caption{Evolution of normalised kinetic energy with normalised time.}
    \label{fig:Normalised_Energy_Time_Slope}
  \end{subfigure}
  \caption{Evolution of kinetic energy with time (MPM) for a pile subjected to 
  gradient input velocities.}
  \label{fig:Energy_Time_Slope}
\end{figure}

@@

\begin{figure}[tbhp]
  \centering
  \begin{subfigure}[b]{0.9\textwidth}
    \includegraphics[width=\textwidth]{figs/Normalised_KEx_Slope}
    \caption{Evolution of normalised horizontal kinetic energy with time.}
    \label{fig:Normalised_KEx_Slope}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{0.9\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/Normalised_KEy_Slope}
    \caption{Evolution of normalised vertical kinetic energy with time.}
    \label{fig:Normalised_KEy_Slope}
  \end{subfigure}
  \caption{Evolution of vertical and horizontal kinetic energy with time (MPM) 
  for a pile subjected to gradient input velocities.}
  \label{fig:Normalised_KEx_KEy_Slope}
\end{figure}

@@

\begin{figure}[tbhp]
  \centering
  \begin{subfigure}[b]{0.92\textwidth}
    \includegraphics[width=\textwidth]{figs/tp_tau_mgd}
    \caption{Power law evolution of $t'_f$ and $\tau$ as a function of kinetic 
    energy $E'_{kx0}$.}
    \label{fig:tp_tau_mgd}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{0.92\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/tp_tau}
    \caption{Linear relationship between decay time and run-out time after the 
    transient as a function of the normalised kinetic energy $E_{kx0}$.}
    \label{fig:tp_tau}
  \end{subfigure}
  \caption{Decay time and run-out time as a function of the normalised kinetic 
  energy $E_{kx0}$.}
  \label{fig:tp_Tau}
\end{figure}

@@

\begin{figure}[tbhp]
  \centering
  \begin{subfigure}[b]{0.9\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/runout_fric_slope}
    \caption{Effect of friction on the run-out distance}
    \label{fig:runout_fric_slope}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{0.9\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/time_fric_slope}
    \caption{Effect of friction on the duration of run-out.}
    \label{fig:time_fric_slope}
  \end{subfigure}
  \caption{MPM simulations of effect of friction on the run-out behaviour of 
  slopes subjected to horizontal excitation.}
  \label{fig:fric_slope}
\end{figure}

>>>>

The DEM sample was composed of $\sim13000$ disks with a uniform distribution of 
diameters by volume fractions ($d_{max} = 1.5 d_{min}$). The mean grain 
diameter and mass are $d\simeq 2.455 $~\si{\mm} and $m\simeq 0.0123$~\si{\kg}, 
respectively. The grains are first poured uniformly into a rectangular box of 
given width and then the right-hand side wall is shifted far to the right 
to allow the grains to spread. A stable granular slope of 13.2~\si{\degree} is 
obtained when all grains come to rest; see~\cref{fig:slope_configuration}. This 
procedure leads to a mean packing fraction $\simeq 0.82$. Soil grains with a 
mean density of 2600~\si{\kg\per\m\cubed} and internal friction coefficient of 
0.4 between grains are considered. For grain-scale simulations, classical DEM
approach was used.

\begin{figure}[tbhp]
  \centering
  \begin{subfigure}[b]{0.47\textwidth}
    \centering
    \includegraphics[width=0.95\textwidth]{figs/slope_configuration}
    \caption{DEM}
    \label{fig:dem_slope_configuration}
  \end{subfigure}
  \begin{subfigure}[b]{0.47\textwidth}
    \centering
    \includegraphics[width=0.95\textwidth]{figs/mpm_velocity_slope}
    \caption{MPM}
    \label{fig:mpm_slope_setup}
  \end{subfigure}
  \caption{Initial geometry and dimensions of the pile subjected to a horizontal 
    velocity of 50J.}
  \label{fig:slope_configuration}
\end{figure}

@@

@@

@@

\Cref{fig:mpm_gradvelocity_200j} shows the initial evolution of the 
granular slope subjected to an initial horizontal energy $E_0 = 61$ (in 
dimensionless units) using MPM. As the granular slope is sheared along the 
bottom, the shear propagates to the top leaving a cavity in the vicinity of the 
left wall. This cavity gets partially filled as the granular mass at the top 
collapse behind the flowing mass due to inertia. Similar behaviour is observed 
during the initial stages of the flow evolution using CD technique 
(\cref{fig:Gradient_Slope_CD_200J}). Due to inertia, the grains at the top 
of the granular heap roll down to fill the cavity, while the pile continues to 
spread.

\begin{figure}[tbph]
  \centering
  \begin{subfigure}[b]{0.47\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/mpm_gradvelocity_200j}
    \caption{MPM simulation of the initial stages of granular pile subjected to a 
    gradient horizontal energy.}
    \label{fig:mpm_gradvelocity_200j}
  \end{subfigure}
  \begin{subfigure}[b]{0.47\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/Gradient_Slope_CD_200J}
    \caption{CD simulation of the initial stages of granular pile subjected to a 
    gradient horizontal energy.~\citep{Mutabaruka2013}.}
    \label{fig:Gradient_Slope_CD_200J}
  \end{subfigure}
  \label{fig:gradvelocity_200j}
\end{figure}

@@

\begin{figure}[tbph]
  \centering
  \begin{subfigure}[b]{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/Runout_Eo_MPM_CD_DEM}
    \caption{Run-out distance as a function of normalised input kinetic energy.}
    \label{fig:Runout_Eo_MPM_CD_DEM}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/Tf_vs_Eo_Slope}
    \caption{Duration of run-out as a function of normalised input kinetic energy.}
    \label{fig:Tf_vs_Eo_Slope}
  \end{subfigure}
  \caption{Evolution of run-out and time as a function of the normalised input 
  energy for a pile subjected a gradient horizontal energy.}
  \label{fig:Slope}
\end{figure}

@@

\begin{figure}[tbhp]
  \centering
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figs/Energy_Slope}
    \caption{Evolution of total kinetic energy with time.}
    \label{fig:energy_slope}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/Normalised_Energy_Time_Slope}
    \caption{Evolution of normalised kinetic energy with normalised time.}
    \label{fig:Normalised_Energy_Time_Slope}
  \end{subfigure}
  \caption{Evolution of kinetic energy with time (MPM) for a pile subjected to 
  gradient input velocities.}
  \label{fig:Energy_Time_Slope}
\end{figure}

@@

\begin{figure}[tbhp]
  \centering
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figs/Normalised_KEx_Slope}
    \caption{Evolution of normalised horizontal kinetic energy with time.}
    \label{fig:Normalised_KEx_Slope}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/Normalised_KEy_Slope}
    \caption{Evolution of normalised vertical kinetic energy with time.}
    \label{fig:Normalised_KEy_Slope}
  \end{subfigure}
  \caption{Evolution of vertical and horizontal kinetic energy with time (MPM) 
  for a pile subjected to gradient input velocities.}
  \label{fig:Normalised_KEx_KEy_Slope}
\end{figure}

@@

\begin{figure}[tbhp]
  \centering
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{figs/tp_tau_mgd}
    \caption{Power law evolution of $t'_f$ and $\tau$ as a function of kinetic 
    energy $E'_{kx0}$.}
    \label{fig:tp_tau_mgd}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/tp_tau}
    \caption{Linear relationship between decay time and run-out time after the 
    transient as a function of the normalised kinetic energy $E_{kx0}$.}
    \label{fig:tp_tau}
  \end{subfigure}
  \caption{Decay time and run-out time as a function of the normalised kinetic 
  energy $E_{kx0}$.}
  \label{fig:tp_Tau}
\end{figure}

@@

\begin{figure}[tbhp]
  \centering
  \begin{subfigure}[b]{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/runout_fric_slope}
    \caption{Effect of friction on the run-out distance}
    \label{fig:runout_fric_slope}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/time_fric_slope}
    \caption{Effect of friction on the duration of run-out.}
    \label{fig:time_fric_slope}
  \end{subfigure}
  \caption{MPM simulations of effect of friction on the run-out behaviour of 
  slopes subjected to horizontal excitation.}
  \label{fig:fric_slope}
\end{figure}
